# Package Use

```{r}
#| include: FALSE

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.asp = 0.618,
  warning = FALSE,
  messahe = FALSE
)
```

```{r}
#| label: setup

library(duke)
library(palmerpenguins)
library(ggmosaic)
library(ggplot2)
library(magrittr)
library(dplyr)
attach(penguins)
```

This vignette aims to comprehensively demonstrate the use and functionality of the package **duke**.
**duke**is fully integrated with the **ggplot2** and allows for the creation of Duke official branded visualizations that are color blind friendly.

```{r}
#| include: false

# Dumby Dataset for Increased Categories
Student <- c("Jack", "Annie", "Paul", "Aidan", "Jake", "Josh", "Grace", "Suzy", "Beth", "Taylor", "Tanner", "Lisa", "Jimmy", "Larry", "Patricia", "Laura", "Yasmin", "Tim")
Grade <- c("A+", "B", "A+", "C", "D", "A+", "E", "C", "B-", "B-", "D", "A-", "B+", "A-", "A-", "D", "B", "E")

dumby_df <- data.frame(Student, Grade)
```

For this purposes of this vignette, we will use the palmerspenguins package, which provides a simple dataset on Antarctic penguins and their characteristics: `penguins` [@horstb].
The dataset has eight variables - some numeric and some categorical - and 344 observations, each representing a unique penguin.

## Scatter Plot - Continuous Color

```{r}
#| label: scatter-c

scatterplot_c <- ggplot(
  penguins,
  aes(x = bill_length_mm, y = bill_depth_mm)
) +
  geom_point(aes(color = body_mass_g)) +
  labs(
    title = "Bill Length vs. Bill Depth",
    x = "Bill Length (mm)", y = "Bill Depth (mm)"
  )

scatterplot_c +
  scale_duke_continuous() +
  theme_duke()
scatterplot_c +
  geom_point(shape = species) +
  scale_duke_continuous() +
  theme_duke()
```

## Scatter Plot - Discrete Color

```{r}
#| label: scatter-d

scatterplot_d <- ggplot(
  penguins,
  aes(x = bill_length_mm, y = bill_depth_mm, color = species)
) +
  geom_point(size = 3) +
  labs(
    title = "Bill Length vs. Bill Depth",
    caption = "(Colors used) \n Duke Royal Blue, Eno, Copper",
    x = "Bill Length (mm)",
    y = "Bill Depth (mm)"
  )

scatterplot_d +
  theme_duke() +
  scale_duke_color_discrete()
scatterplot_d +
  facet_wrap(~species) +
  theme_duke() +
  scale_duke_color_discrete()
```

## Bar Plot

```{r}
#| label: bar

barplot <-
  ggplot(penguins, aes(x = species, fill = species)) +
  geom_bar() +
  labs(
    title = "Distribution of Penguin Species",
    caption = "(Colors from left to right) \n Duke Royal Blue, Duke Navy Blue, Copper",
    x = "Species",
    y = "Count"
  )

barplot +
  scale_duke_fill_discrete() +
  theme_duke()
m_penguins <- penguins %>%
  dplyr::filter(sex == "male")

barplot2 <- ggplot(m_penguins, aes(x = sex, fill = sex)) +
  geom_bar()

barplot2 +
  scale_duke_fill_discrete() +
  theme_duke()

# 8-category plot
barplot8 <- ggplot(dumby_df, aes(x = Grade, fill = Grade)) +
  geom_bar()

barplot8 +
  scale_duke_fill_discrete() +
  theme_duke()

# 7-category plot
barplot7 <- dumby_df %>%
  slice(-13) %>%
  ggplot(aes(x = Grade, fill = Grade)) +
  geom_bar() +
  scale_duke_fill_discrete() +
  theme_duke()

# 6-category plot
barplot6 <- dumby_df %>%
  slice(-c(9, 10, 13)) %>%
  ggplot(aes(x = Grade, fill = Grade)) +
  geom_bar() +
  scale_duke_fill_discrete() +
  theme_duke()

# 5-category plot
barplot5 <- dumby_df %>%
  slice(-c(9, 10, 13, 7, 18)) %>%
  ggplot(aes(x = Grade, fill = Grade)) +
  geom_bar() +
  scale_duke_fill_discrete() +
  theme_duke()

# 4-category plot
barplot4 <- dumby_df %>%
  slice(-c(9, 10, 13, 7, 18, 4, 8)) %>%
  ggplot(aes(x = Grade, fill = Grade)) +
  geom_bar() +
  scale_duke_fill_discrete() +
  theme_duke()
```

## Histogram

```{r}
#| label: hist

histplot <- ggplot(penguins, aes(body_mass_g)) +
  geom_histogram(aes(fill = species), alpha = 0.8) +
  labs(title = "Distribution of Penguin Body Mass", 
       caption = "(Colors used) \n Duke Royal Blue, Duke Navy Blue, Copper", 
       x = "Body Mass (g)", 
       y = "Count")

histplot +
  scale_duke_fill_discrete() +
  theme_duke()
```

## Box Plot

```{r}
#| label: box

boxplot <- ggplot(penguins, aes(sex, body_mass_g)) +
  geom_boxplot() +
  labs(
    title = "Comparison of Body Mass By Sex", 
    x = "Sex", 
    y = "Body Mass (g)"
  )

boxplot +
  theme_duke()
```

## Density Plot

```{r}
#| label: density

densityplot <- ggplot(penguins, aes(bill_depth_mm)) +
  geom_density(aes(fill = species)) +
  labs(
    title = "Density of Penguin Bill Depth",
    x = "Bill Depth (mm)",
    y = "Densiy"
  )

densityplot +
  scale_duke_fill_discrete() +
  theme_duke()
```

## Jitter Plot - Discrete Color

```{r}
#| label: jitter=d

jitterplot_d <- ggplot(penguins, aes(as.factor(year), body_mass_g)) +
  geom_jitter(aes(color = species)) +
  labs(
    title = "Comparison of Body Mass By Year", 
    x = "Year", 
    y = "Body Mass (g)"
  )

jitterplot_d +
  scale_duke_color_discrete() +
  theme_duke()
```

## Jitter Plot - Continuous Color

```{r}
#| label: jitter-c

jitterplot_c <- ggplot(penguins, aes(as.factor(year), body_mass_g)) +
  geom_jitter(aes(color = bill_length_mm)) +
  labs(
    title = "Comparison of Body Mass By Year", 
    x = "Year", 
    y = "Body Mass (g)"
  )

jitterplot_c +
  scale_duke_continuous() +
  theme_duke()
```

## Mosaic Plot

```{r}
#| label: mosaic plot

mosaicplot <- ggplot(data = penguins) +
  ggmosaic::geom_mosaic(aes(x = ggmosaic::product(sex, island), fill = sex)) +
  labs(
    title = "Penguin Sex vs. Island",
    x = "Island",
    y = "Penguin Sex",
    caption = "(Color from top to bottom) \n Granite, Duke Navy Blue, Duke Royal Blue"
  )

mosaicplot +
  scale_duke_fill_discrete() +
  theme_duke()
```

## Label Plot

```{r}
#| label: label

labelplot <- ggplot(
    penguins,
    aes(bill_length_mm, bill_depth_mm,
      fill = island
    )
  ) +
  geom_label(aes(label = island)) +
  labs(
    title = "Bill Length/Depth by Island",
    x = "Bill Length (mm)",
    y = "Bill Depth (mm)"
  )

labelplot +
  scale_duke_fill_discrete() +
  theme_duke()
```

## Quantile Plot

```{r}
#| label: quantile
#| warning: FALSE
#| message: FALSE
quantileplot <-
  ggplot(
    penguins,
    aes(bill_length_mm, bill_depth_mm, color = species)
  ) +
  geom_quantile() +
  labs(title = "Bill Length/Depth Quantiles", 
                x = "Bill Length (mm)", 
                y = "Bill Depth (mm)")

quantileplot +
  scale_duke_color_discrete() +
  theme_duke()
```

## Area Plot

```{r}
#| label: area

areaplot <- ggplot(penguins, aes(body_mass_g, fill = species)) +
  geom_area(stat = "bin") +
  labs(
    title = "Area of Body Mass Index", 
    x = "Body Mass (g)"
  )

areaplot +
  scale_duke_fill_discrete() +
  theme_duke()
```

## Dot Plot

```{r}
#| label: dotplot

dotplot <- ggplot(penguins, aes(body_mass_g)) +
  geom_dotplot(aes(fill = species)) +
  labs(
    title = "Dotplot of Body Mass (g)", 
    x = "Body Mass (g)"
  )

dotplot +
  scale_duke_fill_discrete() +
  theme_duke()
```

## Frequency Polynomial Plot

```{r}
#| label: freqplot

freqplot <-
  ggplot(penguins, aes(body_mass_g)) +
  geom_freqpoly(aes(color = species)) +
  labs(title = "Frequency Polynomial Plot of Body Mass (g)", 
                x = "Body Mass (g)")

freqplot +
  scale_duke_color_discrete() +
  theme_duke()
```

## Column Plot

```{r}
#| label: colplot

colplot <- ggplot(penguins, aes(species, body_mass_g, color = species)) +
  geom_col() +
  labs(
    title = "Body Mass (g) By Species", 
    x = "Species", 
    y = "Body Mass (g)"
  )

colplot +
  scale_duke_color_discrete() +
  theme_duke()
```

## Violin Plot

```{r}
#| label: violin plot

violinplot <- ggplot(penguins, aes(species, body_mass_g, fill = species)) +
  geom_violin(scale = "area") +
  labs(
    title = "Body Mass (g) By Species", 
    x = "Species", 
    y = "Body Mass (g)"
  )

violinplot +
  scale_duke_fill_discrete() +
  theme_duke()
```

## Count Plot

```{r}
#| label: count plot

countplot <- ggplot(penguins, aes(species, island, color = species)) +
  geom_count() +
  labs(
    title = "Species Count by Island",
    x = "Species", 
    y = "Island"
  )

countplot +
  scale_duke_color_discrete() +
  theme_duke()
```

## Step Plot

```{r}
#| label: step plot

stepplot <- ggplot(
  penguins, 
  aes(as.factor(year), body_mass_g, color = species)
  ) +
  geom_step() +
  labs(
    itle = "Body Mass (g) By Year", 
    x = "Year", 
    y = "Body Mass (g)"
  )

stepplot +
  scale_duke_color_discrete() +
  theme_duke()
```
