# Package Use

```{r}
#| include: FALSE
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.asp = .618
)
```

```{r}
#| label: setup
#| warning: FALSE
#| message: FALSE
library(duke)
library(palmerpenguins)
library(ggmosaic)
library(ggplot2)
library(magrittr)
library(dplyr)
attach(penguins)
```

This vignette aims to comprehensively demonstrate the use and functionality of the package **duke**.
**duke**is fully integrated with the **ggplot2** and allows for the creation of Duke official branded visualizations that are color blind friendly.

```{r}
#| include: false
# Dumby Dataset for Increased Categories
Student <- c("Jack", "Annie", "Paul", "Aidan", "Jake", "Josh", "Grace", "Suzy", "Beth", "Taylor", "Tanner", "Lisa", "Jimmy", "Larry", "Patricia", "Laura", "Yasmin", "Tim")
Grade <- c("A+", "B", "A+", "C", "D", "A+", "E", "C", "B-", "B-", "D", "A-", "B+", "A-", "A-", "D", "B", "E")

dumby_df <- data.frame(Student, Grade)
```

For this purposes of this vignette, we will use the palmerspenguins package, which provides a simple dataset on Antarctic penguins and their characteristics [@horstb].
The data has 8 variables - both numeric and categorical - and 344 observations, each representing a unique penguin.

## Scatter Plot - Continuous Color

```{r}
#| label: scatter c
#| warning: FALSE
#| message: FALSE
scatterplot_c <- ggplot(palmerpenguins::penguins, 
                        aes(x = bill_length_mm, y = bill_depth_mm)) +
                  geom_point(aes(color = body_mass_g)) +
                  labs(title = "Bill Length vs. Bill Depth", 
                       x = "Bill Length (mm)", y = "Bill Depth (mm)")

scatterplot_c +
  scale_duke_continuous() +
  theme_duke()
scatterplot_c +
  geom_point(shape = species) +
  scale_duke_continuous() +
  theme_duke()
```

## Scatter Plot - Discrete Color

```{r}
#| label: scatter d
#| warning: FALSE
#| message: FALSE
scatterplot_d <-
  ggplot(palmerpenguins::penguins,
         aes(x = bill_length_mm, y = bill_depth_mm, color = species)) +
  geom_point(size = 3) +
  labs(
    title = "Bill Length vs. Bill Depth",
    caption = "(Colors used) \n Duke Royal Blue, Eno, Copper",
    x = "Bill Length (mm)",
    y = "Bill Depth (mm)"
  )

scatterplot_d +
  theme_duke() +
  scale_duke_color_discrete()
scatterplot_d +
  facet_wrap( ~ species) +
  theme_duke() +
  scale_duke_color_discrete()
```

## Bar Plot

```{r}
#| label: bar
#| warning: FALSE
#| message: FALSE
barplot <-
  ggplot(palmerpenguins::penguins, aes(x = species, fill = species)) +
  geom_bar() +
  labs(
    title = "Distribution of Penguin Species",
    caption = "(Colors from left to right) \n Duke Royal Blue, Duke Navy Blue, Copper",
    x = "Species",
    y = "Count"
  )

barplot +
  scale_duke_fill_discrete() +
  theme_duke()
m_penguins <- penguins %>%
  dplyr::filter(sex == "male")

barplot2 <- ggplot(m_penguins, aes(x = sex, fill = sex)) +
  geom_bar()

barplot2 +
  scale_duke_fill_discrete() +
  theme_duke()
# 8-category plot
barplot8 <- ggplot(dumby_df, aes(x = Grade, fill = Grade)) +
  geom_bar()

barplot8 +
  scale_duke_fill_discrete() +
  theme_duke()
# 7-category plot
barplot7 <- dumby_df %>%
  slice(-13) %>%
  ggplot(aes(x = Grade, fill = Grade)) +
  geom_bar() +
  scale_duke_fill_discrete() +
  theme_duke()
# 6-category plot
barplot6 <- dumby_df %>%
  slice(-c(9, 10, 13)) %>%
  ggplot(aes(x = Grade, fill = Grade)) +
  geom_bar() +
  scale_duke_fill_discrete() +
  theme_duke()
# 5-category plot
barplot5 <- dumby_df %>%
  slice(-c(9, 10, 13, 7, 18)) %>%
  ggplot(aes(x = Grade, fill = Grade)) +
  geom_bar() +
  scale_duke_fill_discrete() +
  theme_duke()
# 4-category plot
barplot4 <- dumby_df %>%
  slice(-c(9, 10, 13, 7, 18, 4, 8)) %>%
  ggplot(aes(x = Grade, fill = Grade)) +
  geom_bar() +
  scale_duke_fill_discrete() +
  theme_duke()
```

## Histogram

```{r}
#| label: hist
#| warning: FALSE
histplot <- ggplot2::ggplot(palmerpenguins::penguins, aes(body_mass_g)) +
  geom_histogram(ggplot2::aes(fill = species), alpha = 0.8) +
  labs(title = "Distribution of Penguin Body Mass", 
       caption = "(Colors used) \n Duke Royal Blue, Duke Navy Blue, Copper", 
       x = "Body Mass (g)", 
       y = "Count")

histplot +
  scale_duke_fill_discrete() +
  theme_duke()
```

## Box Plot

```{r}
#| label: box
#| warning: FALSE
#| message: FALSE
boxplot <- ggplot2::ggplot(palmerpenguins::penguins, ggplot2::aes(sex, body_mass_g)) +
  ggplot2::geom_boxplot() +
  ggplot2::labs(title = "Comparison of Body Mass By Sex", 
                x = "Sex", 
                y = "Body Mass (g)")

boxplot +
  theme_duke()
```

## Density Plot

```{r}
#| label: density
#| warning: FALSE
densityplot <- ggplot2::ggplot(palmerpenguins::penguins, ggplot2::aes(bill_depth_mm)) +
  ggplot2::geom_density(ggplot2::aes(fill = species)) +
  ggplot2::labs(title = "Density of Penguin Bill Depth", 
                x = "Bill Depth (mm)", 
                y = "Densiy")

densityplot +
  scale_duke_fill_discrete() +
  theme_duke()
```

## Jitter Plot - Discrete Color

```{r}
#| label: jitter d
#| warning: FALSE
#| message: FALSE
jitterplot_d <- ggplot2::ggplot(palmerpenguins::penguins, ggplot2::aes(as.factor(year), body_mass_g)) +
  ggplot2::geom_jitter(ggplot2::aes(color = species)) +
  ggplot2::labs(title = "Comparison of Body Mass By Year", 
                x = "Year", 
                y = "Body Mass (g)")

jitterplot_d +
  scale_duke_color_discrete() +
  theme_duke()
```

## Jitter Plot - Continuous Color

```{r}
#| label: jitter c
#| warning: FALSE
#| message: FALSE
jitterplot_c <- ggplot2::ggplot(palmerpenguins::penguins, ggplot2::aes(as.factor(year), body_mass_g)) +
  ggplot2::geom_jitter(ggplot2::aes(color = bill_length_mm)) +
  ggplot2::labs(title = "Comparison of Body Mass By Year", 
                x = "Year", 
                y = "Body Mass (g)")

jitterplot_c +
  scale_duke_continuous() +
  theme_duke()
```

## Mosaic Plot

```{r}
#| label: mosaic plot
#| warning: FALSE
#| message: FALSE
mosaicplot <- ggplot(data = palmerpenguins::penguins) +
  ggmosaic::geom_mosaic(aes(x = ggmosaic::product(sex, island), fill = sex)) +
  labs(title = "Penguin Sex vs. Island", 
       x = "Island", 
       y = "Penguin Sex", 
       caption = "(Color from top to bottom) \n Granite, Duke Navy Blue, Duke Royal Blue")

mosaicplot +
  scale_duke_fill_discrete() +
  theme_duke()
```

## Label Plot

```{r}
#| label: label
#| warning: FALSE
#| message: FALSE
labelplot <-
  ggplot2::ggplot(
    palmerpenguins::penguins,
    ggplot2::aes(bill_length_mm, bill_depth_mm, 
                 fill = island)
  ) +
  ggplot2::geom_label(aes(label = island)) +
  ggplot2::labs(title = "Bill Length/Depth by Island", 
                x = "Bill Length (mm)", 
                y = "Bill Depth (mm)")

labelplot +
  scale_duke_fill_discrete() +
  theme_duke()
```

## Quantile Plot

```{r}
#| label: quantile
#| warning: FALSE
#| message: FALSE
quantileplot <-
  ggplot2::ggplot(
    palmerpenguins::penguins,
    ggplot2::aes(bill_length_mm, bill_depth_mm, color = species)
  ) +
  ggplot2::geom_quantile() +
  ggplot2::labs(title = "Bill Length/Depth Quantiles", 
                x = "Bill Length (mm)", 
                y = "Bill Depth (mm)")

quantileplot +
  scale_duke_color_discrete() +
  theme_duke()
```

## Area Plot

```{r}
#| label: area
#| warning: FALSE
#| message: FALSE
areaplot <-
  ggplot2::ggplot(palmerpenguins::penguins,
                  ggplot2::aes(body_mass_g, fill = species)) +
  ggplot2::geom_area(stat = "bin") +
  ggplot2::labs(title = "Area of Body Mass Index", 
                x = "BMI")

areaplot +
  scale_duke_fill_discrete() +
  theme_duke()
```

## Dot Plot

```{r}
#| label: dotplot
#| warning: FALSE
#| message: FALSE
dotplot <-
  ggplot2::ggplot(palmerpenguins::penguins, ggplot2::aes(body_mass_g)) +
  ggplot2::geom_dotplot(aes(fill = species)) +
  ggplot2::labs(title = "Dotplot of BMI", 
                x = "BMI")

dotplot +
  scale_duke_fill_discrete() +
  theme_duke()
```

## Frequency Polynomial Plot

```{r}
#| label: freqplot
#| warning: FALSE
#| message: FALSE
freqplot <-
  ggplot2::ggplot(palmerpenguins::penguins, ggplot2::aes(body_mass_g)) +
  ggplot2::geom_freqpoly(aes(color = species)) +
  ggplot2::labs(title = "Frequency Polynomial Plot of BMI", 
                x = "BMI")

freqplot +
  scale_duke_color_discrete() +
  theme_duke()
```

## Column Plot

```{r}
#| label: colplot
#| warning: FALSE
#| message: FALSE
colplot <-
  ggplot2::ggplot(palmerpenguins::penguins,
                  ggplot2::aes(species, body_mass_g, color = species)) +
  ggplot2::geom_col() +
  ggplot2::labs(title = "BMI By Species", 
                x = "Species", 
                y = "BMI")

colplot +
  scale_duke_color_discrete() +
  theme_duke()
```

## Violin Plot

```{r}
#| label: violin plot
#| warning: FALSE
#| message: FALSE
violinplot <-
  ggplot2::ggplot(palmerpenguins::penguins,
                  ggplot2::aes(species, body_mass_g, fill = species)) +
  geom_violin(scale = "area") +
  ggplot2::labs(title = "BMI By Species", 
                x = "Species", 
                y = "BMI")

violinplot +
  scale_duke_fill_discrete() +
  theme_duke()
```

## Count Plot

```{r}
#| label: count plot
#| warning: FALSE
#| message: FALSE
countplot <-
  ggplot2::ggplot(palmerpenguins::penguins,
                  ggplot2::aes(species, island, color = species)) +
  geom_count() +
  ggplot2::labs(title = "Species Count by Island", 
                x = "Species", 
                y = "Island")

countplot +
  scale_duke_color_discrete() +
  theme_duke()
```

## Step Plot

```{r}
#| label: step plot
#| warning: FALSE
#| message: FALSE
stepplot <-
  ggplot2::ggplot(palmerpenguins::penguins,
                  ggplot2::aes(as.factor(year), body_mass_g, 
                               color = species)) +
  geom_step() +
  
  ggplot2::labs(title = "BMI By Year", 
                x = "Year", 
                y = "BMI")

stepplot +
  scale_duke_color_discrete() +
  theme_duke()
```
